Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-05-19 document/ms.tex
I can't spell

An impressive number of exoplanets have been discovered and it is of great
scientific interest to study what we can say about the exoplanet occurence
rate function given these discoveries.
Every catalog is censored by non-trivial selection effects and detection
efficiencies, and the measurements of the system properties (period, radius,
\etc) are noisy.
We present a general hierarchical probabilistic framework for making justified
inferences about the occurrence rate of exoplanets taking into account survey
completeness \emph{and observational uncertainties}.
By applying our method to synthetic catalogs, we demonstrate that it produces
more accurate estimates of the underlying occurence rate than the standard
inverse-detection-efficiency weighting procedure.
When applied to an existing catalog of small planet candidates around G dwarfs
(Petigura \etal\ 2013), our method confirms previous measurements of a
flattening radius distribution near Earth's radius.
Although no true extra-solar Earth analog is known, hundreds of planets have
been found around Sun-like stars that are either Earth-sized but on shorter
periods, or else on year-long orbits but somewhat larger.
Under strong assumptions, it has been shown that these populations permit an
uncertain extrapolated estimate of the probability or rate at which Sun-like
stars host true Earth analogs.
We relax these assumptions by modeling the occurence rate as a
multidimensional Gaussian process and find that the rate density of Earth
analogs is about two percent (per natural logarithmic bin in period and
radius) at low signal-to-noise.

@@

% CITATIONS:
% \begin{itemize}
% \item
% \citet{fressin-fp} --- simultaneous occurence rate and false positive rate
% \item
% \citet{morton} --- false positive rate
% \item
% \citet{morton-swift} --- KDE estimator for radius distribution of cool KOIs
% \item
% \citet{howard} --- inv-det-eff for short-period candidates around Sun-like
% stars
% \item
% \citet{dressing} --- inv-det-eff for cool KOIs
% \item
% \citet{dong} --- semi-analytic completeness, likelihood function-ish,
% turn-over at small radii, non-separability
% \item
% \citet{tabachnik} ---- Poisson likelihood
% \item
% \citet{tremaine} --- statistics of multis
% \item
% \citet{youdin} --- Poisson likelihood
% \item
% \citet{catanzarite} --- extrapolation to Eta-Earth assuming complete
% \item
% \citet{traub} --- extrapolation to Eta-Earth assuming complete
% \end{itemize}

NASA's \kepler\ mission has enabled the discovery of thousands of exoplanet
candidates (\dfmtodo{CITE KEPLER}).
Many of these planets orbit Sun-like stars (\citealt{petigura}), where the
definition of Sun-like is given in terms of the star's temperature and surface
gravity.
Given these catalogs, it is interesting to ask what we can say about the
occurence rate of exoplanets as a function of their physical parameters
(period, radius, \etc).
The two main reasons why this is interesting are that \emph{(a)} theories of
planet formation make predictions about the shape of this function
(\dfmtodo{CITE}), and \emph{(b)} it allows us to place probabilistic
constraints on the rate of Earth analogs\footnote{We define, in detail, what
we mean by Earth analog below but, in short, we mean Earth-sized planets on
year-long orbits around Sun-like stars.}.

\citet{petigura} recently published constraints on this function based on
their independent analysis of the \kepler\ light curves for 42,000 Sun-like
stars but the study of exoplanet occurence rates has a long history.
\dfmtodo{Summarize history with lotsa citations.}

A very important component of any study of exoplanet occurence rates is the
treatment of detection efficiency.
Qualitatively, in a transit survey, small planets with long periods are much
harder to detect than large planets orbiting close to their star.
This effect is degenerate with any inferences about the rate function and it
can be hard to constraint quantitatively.
In practice, there are three methods for taking this effect into account:
\emph{(a)} making conservative cuts on the candidates and assuming that the
resulting catalog is complete (\citealt{catanzarite, traub}; \dfmtodo{CITE}),
\emph{(b)} asserting an analytic form for the detection efficiency as a
function of approximate signal-to-noise (\citealt{howard, dressing, dong,
fressin-fp, morton-swift}; \dfmtodo{CITE}), and \emph{(c)} determining the
detection efficiency empirically by injecting synthetic signals into the raw
data and testing recovery rates (\citealt{petigura-a, petigura}).

In this \paper, we develop a flexible framework for probabilistic inference of
exoplanet occurence rates that can be applied to incomplete catalogs with
\emph{non-negligible observational uncertainties}.
Our method is a generalization of a previously studied method
(\citealt{hogge}) to account for detection efficiencies.
We run tests on simulated datasets---comparing results with the standard
techniques that neglect observational uncertainties---and apply our method to
a real catalog of small planets transiting Sun-like stars
(\citealt{petigura}).
For the purposes of this \paper, we make some strong assumptions---although we
argue that they are weaker than the implicit assumptions in previous
studies---but none of these are necessary for the applicability of our method.
Specifically, we make the following assumptions
\begin{itemize}

\item the candidates in the catalog are independent Poisson draws from the
censored occurence rate,

@@

\item the true occurence rate function is \emph{smooth}\footnote{We give our
definition of ``smooth'' in more detail below but our model is very flexible
so this is not a strong restriction.}.

@@

In this \paper, we are asking the question: \emph{what constraints can we put
on the occurence rate of exoplanets given all the light curves measured by
\kepler?}
Throughout this \paper, we use the notation $\rate_\ratepars(\entry)$ for the
occurence rate \rate---parameterized by the parameters \ratepars---as a
function of the physical parameters \entry\ (orbital period, planetary radius,
\etc).
In this framework, the occurence rate function can be ``parametric''---for
example, a power law---or a ``non-parametric'' function---such as a histogram
where the bin heights are the parameters \ratepars.
The light curve for a particular target $k$ is called $\data_k$ and the
\emph{true} set of physical parameters for that same object are given by
$\entry_k$.
A catalog provides a set of constraints on the parameters $\{\entry_k\}$
conditioned the set light curves $\{\data_k\}$.

Any inference about the rate function parameters can be formally addressed by
evaluating the \emph{marginalized likelihood function}
\begin{eqnarray}\eqlabel{crazylike}
p(\{\data_k\}\,|\,\ratepars) &=&
    \int p(\{\data_k\}\,|\,\{\entry_k\})
    \,p(\{\entry_k\}\,|\,\ratepars)
    \dd\{\entry_k\}
\end{eqnarray}
where, as we mentioned above, $\{\data_k\}$ is the set of all light curves,
one light curve $\data_k$ per target $k$, \ratepars\ is the vector of
parameters describing the occurence rate function $\rate_\ratepars(\entry)$
and $\entry_k$ is the vector of physical parameters describing the planetary
system (orbital periods, radius ratios, stellar radius, \etc) around target
$k$.
In this equation, our only assumption is that the datasets depend on the
occurence rate of exoplanets only through the catalog $\{\entry_k\}$.
In our case, this assumption qualitatively means that the signals found in the
light curves depend only on the actual planet properties and not on the
distributions from which they are drawn.
It is worth noting that---as we will discuss further below---the catalog only
provides \emph{probabilistic} constraints on $\{\entry_k\}$.

@@

In the inference procedure described above, we only discussed an abstract
occurence rate function.
To make it more concrete, we'll model the catalog as a Poisson draw from the
\emph{observable} rate function $\obs{\rate}_\ratepars$.
This leads to the previously known result (see \citealt{tabachnik,youdin} for
some of the examples from the exoplanet literature)
\begin{eqnarray}\eqlabel{poisson-like}
p(\{\entry_k\}\,|\,\ratepars) &=&
    \exp\left(-\int \obs{\rate}_\ratepars (\entry) \dd\entry\right) \,
    \prod_{k=1}^K \obs{\rate}_\ratepars (\entry_k)\quad.
\end{eqnarray}
In this equation, the integral in the normalization term is the expected
number of observable exoplanets in the sample.
A further assumption that we have made in \eq{poisson-like} is that each
object is an independent sample from the rate function.
This assumption---though it is often made---cannot be correct when we consider
systems of multiple planets.
It is possible to relax this assumption (see \citealt{tremaine} for a related
technique) but this generalization is beyond the scope of the current \paper.

The main thing to note here is that $\obs{\rate}_\ratepars$ is the rate of
exoplanets that you would expect to observe taking into account the geometric
transit probability and any other detection efficiencies.
In practice, it probably makes sense to model the observable rate as
\begin{eqnarray}
\obs{\rate}_\ratepars(\entry) &=&
    \completeness(\entry)\,\rate_\ratepars(\entry)
\end{eqnarray}
where $\completeness(\entry)$ is the detection efficiency at \entry\ and
$\rate_\ratepars(\entry)$ is the object that we want to infer: the true
occurence rate.
Up to this point, we haven't discussed any specific functional form for
$\rate_\ratepars(\entry)$ and all of this derivation is equally applicable
whether we model the occurence rate as (for example) a broken power law or
a histogram.

@@

Now, we can substitute \eq{poisson-like} into \eq{simplemarglike} and apply
the importance sampling approximation to derive the following expression for
the marginalized likelihood
\begin{eqnarray}\eqlabel{money}
\frac{p(\{\data_k\}\,|\,\ratepars)}{p(\{\data_k\}\,|\,\interim)} &\approx&
    \exp\left(-\int \obs{\rate}_\ratepars (\entry) \dd\entry\right) \,
    \prod_{k=1}^K
    \frac{1}{N_k} \sum_{n=1}^{N_k}
    \frac{\obs{\rate}_\ratepars (\entry_k^{(n)})}
         {p(\entry_k^{(n)}\,|\,\interim)} \quad.
\end{eqnarray}
In this equation, we're making the further assumption that the catalog treated
the objects independently.
In other words, we have per-object posterior samples
\begin{eqnarray}
\entry_k^{(n)} &\sim& p(\entry_k\,|\,\data_k,\,\interim) \quad.
\end{eqnarray}
\Eq{money} is the \emph{money equation} for our method.
It lets us efficiently compute the \emph{marginalized likelihood of the entire
set of light curves for a particular occurence rate function}.

@@

If we model the occurence rate as a histogram with $J$ fixed bin volumes
$\binarea_j$
\begin{eqnarray}
\rate_\ratepars (\entry) &=& \left\{\begin{array}{ll}
\ratepar_1 & \entry \in \binarea_1,\\
\ratepar_2 & \entry \in \binarea_2,\\
\cdots \\
\ratepar_J & \entry \in \binarea_J,\\
0 & \mathrm{otherwise}
\end{array}\right.
\end{eqnarray}
then \eq{poisson-like} becomes
\begin{eqnarray}
\ln p(\{\entry_k\}\,|\,\ratepars) &=&
    \sum_{k=1}^K \sum_{j=1}^J \mathbf{1}[\entry_k \in
        \binarea_j]\,\ln\completeness(\entry_k)\,\ratepar_j
    -\sum_{j=1}^J\ratepar_j\,\int_{\binarea_j} \completeness(\entry)\dd\entry
\end{eqnarray}
where the indicator function $\mathbf{1}[\cdot]$ is one if $\cdot$ is true and
zero otherwise.
If we take the gradient of this function with respect to \ratepars\ and set it
equal to zero, we find the maximum likelihood result
\begin{eqnarray}
\ratepar_j &=& \frac{N_j}{\int_{\binarea_j} \completeness(\entry)\dd\entry}
\end{eqnarray}
where $N_j$ is the number of objects that fall within the bin $j$.
Since it is the maximum likelihood solution, this result is the minimum
variance estimator for the occurence rate in the limit of negligible catalog
uncertainties.
This is \emph{almost} the same as the inverse-detection-efficiency result and
in various limiting cases, they would provide the same result.
\dfmtodo{Finish explaining why this result is better under the specified
assumptions.}

\paragraph{Occurence rate parameterization \& priors}
For the remainder of this \paper, we also model the occurence rate as a
two-dimensional histogram with fixed logarithmic bins in period and radius.
When we include observational uncertainties---using \eq{money}---the maximum
likelihood result is no longer analytic.
Therefore, if we want to compute the ``best-fit'' occurence rate, we can use a
standard non-linear optimization algorithm.
In the regions of parameter space that we tend to care about, the completeness
is low and there are only a few observations with large uncertainties.
In this case, we're especially interested in probabilistic constraints on the
occurence rate; not just the best-fit model.
To do this, we apply a prior\footnote{This is a slight misnomer because
we also we also marginalize over the parameters of this model.} $p(\ratepars)$
on the occurence rate parameters and generate samples from the posterior
probability
\begin{eqnarray}\eqlabel{posterior}
p(\ratepars\,|\,\{\data_k\}) &\propto&
    p(\ratepars)\,p(\{\data_k\}\,|\,\ratepars)
\end{eqnarray}
using Markov chain Monte Carlo (MCMC).

Given our model of the rate function as a collection of bins, we choose to
model the prior as a Gaussian process (\citealt{gp,gibson-gp}) on the
logarithmic bin heights.
This model encodes our prior belief that, on the grid scale that we consider,
the occurence rate should be smooth but it is otherwise very flexible about
the form of the function.
Mathematically, the prior is written as
\begin{eqnarray}
p(\ratepars) &=& p(\ratepars\,|\,\mean,\,\smooth) \\
&=& \mathcal{N} \left[\ln\ratepars;\,\mean\,\bvec{I},\,
\cov(\{\binarea_j\},\,\smoothpars)\right]
\end{eqnarray}
where $\mathcal{N}(\cdot;\,\mean\,\bvec{I},\,\cov)$ is a $J$-dimensional
Gaussian with a constant mean \mean\ and covariance matrix \cov\ that depends
on the bin locations $\{\binarea_j\}$ and a set of hyperparameters
$\smoothpars = (\smooth_0,\,\smooth_\period,\,\smooth_\radius)$.
The covariance function that we use is an anisotropic, axis-aligned
exponential-squared kernel so elements of the matrix are
\begin{eqnarray}
\cov_{ij} &=& \smooth_0\,\exp\left(-\frac{1}{2}\,
    [\binarea_i-\binarea_j]^\mathrm{T}\,\Sigma^{-1}\,[\binarea_i-\binarea_j]
\right)
\end{eqnarray}
where $\Sigma^{-1}$ is the diagonal matrix
\begin{eqnarray}
\Sigma^{-1} &=& \left(\begin{array}{cc}
1/\smooth_\period & 0 \\
0 & 1/\smooth_\radius
\end{array}\right) \quad.
\end{eqnarray}

@@

In order to get a feeling for the constraints provided by our method and to
explore any biases introduced by ignoring the observational uncertainties, we
start by ``observing'' a few synthetic catalogs from qualitatively different
known occurence rate functions.
For each of these simulations, we'll take the completeness function computed
by \citet{petigura} as given.
In general, \eq{poisson-like} can be sampled using a procedure called thinning
(\citealt{poisson}) but for our purposes, we'll simply consider a piecewise
constant rate function evaluated on a fine grid in log-period and log-radius.
For this discrete function, the generative procedure is simple;
\begin{enumerate}
{\item loop over each grid cell $i$,}
{\item draw Poisson random integer $K_i\sim\mathrm{Poissson}(\obs{\rate}_i)$
with the observable rate evaluated in the cell, and}
{\item distribute $K_i$ catalog entries in the cell randomly.}
\end{enumerate}
We then choose fractional observational uncertainties on the radii from the
\citet{petigura} catalog and apply them to the true catalog as Gaussian noise.

We generate synthetic catalogs from two qualitatively different rate
functions.
Both distributions are generated by a separable model
\begin{eqnarray}
\rate_\ratepars (\ln\period,\,\ln\radius) &=&
    \rate_\ratepars^{(\period)}(\ln\period)\,
    \rate_\ratepars^{(\radius)}(\ln\radius)
\end{eqnarray}
but fit using the full general model.
The first catalog---\modela--- is generated assuming a smooth occurence
surface where both distributions are broken power laws.
The second---\modelb---is designed to be exactly the distribution inferred by
\citet{petigura} in the range that they considered and then smoothly
extrapolated outside that range.
The catalogs generated from these two models are shown in \fig{smooth-results}
and \fig{simulation-results}, respectively and the data are available online
at \dfmtodo{ADD SOME URL}.

For each catalog, we directly apply the inverse-detection-efficiency procedure
as implemented by \citealt{petigura}\footnote{Our implementation reproduces
their results when applied to the published catalog.} and our probabilistic
method, marginalizing over the hyperparameters of the Gaussian process
regularization.
\Fig{smooth-results} and \fig{simulation-results} show the results of this
analysis in both cases.
In particular, the side panels compare the marginalized occurence rate
functions in period and radius to the true functions that were used to
generate the catalogs.
\Fig{smooth-results} shows that even if the true occurence rate is a smooth
function, the rate inferred by the inverse-detection-efficiency method can
appear to have sharp features.
In this first example---where the true distribution is well described by our
Gaussian process model---the probabilistic inference of the occurence rate is
both more precise and accurate.

In the second example, the true rate function includes a sharp feature chosen
to reproduce the result published by \citet{petigura}.
In this case, \fig{simulation-results} shows that the probabilistic
constraints on the occurence rate are less precise \emph{but more accurate}
than results using the inverse-detection-efficiency method.
At this point, the distributions inferred by these two different techniques
don't seem drastically inconsistent but the difference are most obvious as the
survey completeness goes to zero: exactly the region of interest.

\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{figures/smooth/results.pdf}
\end{center}
\caption{%
{\bf Simulated data}.
Inferences about the occurence rate function based on the simulated catalog
\modela.
\emph{Center:} the points with error bars show the exoplanet candidates in the
simulated incomplete catalog, the contours show the survey completeness
function (\citealt{petigura}), and the grayscale shows the median posterior
occurence surface.
\emph{Top and left:} the red dashed line shows the \emph{true} distribution
that was used to generate the catalog, the points with error bars show the
results of the inverse-detection-efficiency procedure, and the histograms are
posterior samples from the marginalized occurence rate as inferred by our
method.
\figlabel{smooth-results}}
\end{figure}

@@

Having developed this probabilistic framework for occurence rate inferences
and demonstrating that it produces reasonable results when applied to
simulated datasets, we now turn to real data.
As described in \sect{data}, we will use the catalog of small exoplanet
candidates orbiting Sun-like stars published by \citet{petigura}.
This is a great test case because those authors empirically measured the
detection efficiency of their pipeline as a function of the parameters of
interest.

>>>>

An impressive number of exoplanets have been discovered and it is of great
scientific interest to study what we can say about the exoplanet occurrence
rate function given these discoveries.
Every catalog is censored by non-trivial selection effects and detection
efficiencies, and the measurements of the system properties (period, radius,
\etc) are noisy.
We present a general hierarchical probabilistic framework for making justified
inferences about the occurrence rate of exoplanets taking into account survey
completeness \emph{and observational uncertainties}.
By applying our method to synthetic catalogs, we demonstrate that it produces
more accurate estimates of the underlying occurrence rate than the standard
inverse-detection-efficiency weighting procedure.
When applied to an existing catalog of small planet candidates around G dwarfs
(Petigura \etal\ 2013), our method confirms previous measurements of a
flattening radius distribution near Earth's radius.
Although no true extra-solar Earth analog is known, hundreds of planets have
been found around Sun-like stars that are either Earth-sized but on shorter
periods, or else on year-long orbits but somewhat larger.
Under strong assumptions, it has been shown that these populations permit an
uncertain extrapolated estimate of the probability or rate at which Sun-like
stars host true Earth analogs.
We relax these assumptions by modeling the occurrence rate as a
multidimensional Gaussian process and find that the rate density of Earth
analogs is about two percent (per natural logarithmic bin in period and
radius) at low signal-to-noise.

@@

% CITATIONS:
% \begin{itemize}
% \item
% \citet{fressin-fp} --- simultaneous occurrence rate and false positive rate
% \item
% \citet{morton} --- false positive rate
% \item
% \citet{morton-swift} --- KDE estimator for radius distribution of cool KOIs
% \item
% \citet{howard} --- inv-det-eff for short-period candidates around Sun-like
% stars
% \item
% \citet{dressing} --- inv-det-eff for cool KOIs
% \item
% \citet{dong} --- semi-analytic completeness, likelihood function-ish,
% turn-over at small radii, non-separability
% \item
% \citet{tabachnik} ---- Poisson likelihood
% \item
% \citet{tremaine} --- statistics of multis
% \item
% \citet{youdin} --- Poisson likelihood
% \item
% \citet{catanzarite} --- extrapolation to Eta-Earth assuming complete
% \item
% \citet{traub} --- extrapolation to Eta-Earth assuming complete
% \end{itemize}

NASA's \kepler\ mission has enabled the discovery of thousands of exoplanet
candidates (\dfmtodo{CITE KEPLER}).
Many of these planets orbit Sun-like stars (\citealt{petigura}), where the
definition of Sun-like is given in terms of the star's temperature and surface
gravity.
Given these catalogs, it is interesting to ask what we can say about the
occurrence rate of exoplanets as a function of their physical parameters
(period, radius, \etc).
The two main reasons why this is interesting are that \emph{(a)} theories of
planet formation make predictions about the shape of this function
(\dfmtodo{CITE}), and \emph{(b)} it allows us to place probabilistic
constraints on the rate of Earth analogs\footnote{We define, in detail, what
we mean by Earth analog below but, in short, we mean Earth-sized planets on
year-long orbits around Sun-like stars.}.

\citet{petigura} recently published constraints on this function based on
their independent analysis of the \kepler\ light curves for 42,000 Sun-like
stars but the study of exoplanet occurrence rates has a long history.
\dfmtodo{Summarize history with lotsa citations.}

A very important component of any study of exoplanet occurrence rates is the
treatment of detection efficiency.
Qualitatively, in a transit survey, small planets with long periods are much
harder to detect than large planets orbiting close to their star.
This effect is degenerate with any inferences about the rate function and it
can be hard to constraint quantitatively.
In practice, there are three methods for taking this effect into account:
\emph{(a)} making conservative cuts on the candidates and assuming that the
resulting catalog is complete (\citealt{catanzarite, traub}; \dfmtodo{CITE}),
\emph{(b)} asserting an analytic form for the detection efficiency as a
function of approximate signal-to-noise (\citealt{howard, dressing, dong,
fressin-fp, morton-swift}; \dfmtodo{CITE}), and \emph{(c)} determining the
detection efficiency empirically by injecting synthetic signals into the raw
data and testing recovery rates (\citealt{petigura-a, petigura}).

In this \paper, we develop a flexible framework for probabilistic inference of
exoplanet occurrence rates that can be applied to incomplete catalogs with
\emph{non-negligible observational uncertainties}.
Our method is a generalization of a previously studied method
(\citealt{hogge}) to account for detection efficiencies.
We run tests on simulated datasets---comparing results with the standard
techniques that neglect observational uncertainties---and apply our method to
a real catalog of small planets transiting Sun-like stars
(\citealt{petigura}).
For the purposes of this \paper, we make some strong assumptions---although we
argue that they are weaker than the implicit assumptions in previous
studies---but none of these are necessary for the applicability of our method.
Specifically, we make the following assumptions
\begin{itemize}

\item the candidates in the catalog are independent Poisson draws from the
censored occurrence rate,

@@

\item the true occurrence rate function is \emph{smooth}\footnote{We give our
definition of ``smooth'' in more detail below but our model is very flexible
so this is not a strong restriction.}.

@@

In this \paper, we are asking the question: \emph{what constraints can we put
on the occurrence rate of exoplanets given all the light curves measured by
\kepler?}
Throughout this \paper, we use the notation $\rate_\ratepars(\entry)$ for the
occurrence rate \rate---parameterized by the parameters \ratepars---as a
function of the physical parameters \entry\ (orbital period, planetary radius,
\etc).
In this framework, the occurrence rate function can be ``parametric''---for
example, a power law---or a ``non-parametric'' function---such as a histogram
where the bin heights are the parameters \ratepars.
The light curve for a particular target $k$ is called $\data_k$ and the
\emph{true} set of physical parameters for that same object are given by
$\entry_k$.
A catalog provides a set of constraints on the parameters $\{\entry_k\}$
conditioned the set light curves $\{\data_k\}$.

Any inference about the rate function parameters can be formally addressed by
evaluating the \emph{marginalized likelihood function}
\begin{eqnarray}\eqlabel{crazylike}
p(\{\data_k\}\,|\,\ratepars) &=&
    \int p(\{\data_k\}\,|\,\{\entry_k\})
    \,p(\{\entry_k\}\,|\,\ratepars)
    \dd\{\entry_k\}
\end{eqnarray}
where, as we mentioned above, $\{\data_k\}$ is the set of all light curves,
one light curve $\data_k$ per target $k$, \ratepars\ is the vector of
parameters describing the occurrence rate function $\rate_\ratepars(\entry)$
and $\entry_k$ is the vector of physical parameters describing the planetary
system (orbital periods, radius ratios, stellar radius, \etc) around target
$k$.
In this equation, our only assumption is that the datasets depend on the
occurrence rate of exoplanets only through the catalog $\{\entry_k\}$.
In our case, this assumption qualitatively means that the signals found in the
light curves depend only on the actual planet properties and not on the
distributions from which they are drawn.
It is worth noting that---as we will discuss further below---the catalog only
provides \emph{probabilistic} constraints on $\{\entry_k\}$.

@@

In the inference procedure described above, we only discussed an abstract
occurrence rate function.
To make it more concrete, we'll model the catalog as a Poisson draw from the
\emph{observable} rate function $\obs{\rate}_\ratepars$.
This leads to the previously known result (see \citealt{tabachnik,youdin} for
some of the examples from the exoplanet literature)
\begin{eqnarray}\eqlabel{poisson-like}
p(\{\entry_k\}\,|\,\ratepars) &=&
    \exp\left(-\int \obs{\rate}_\ratepars (\entry) \dd\entry\right) \,
    \prod_{k=1}^K \obs{\rate}_\ratepars (\entry_k)\quad.
\end{eqnarray}
In this equation, the integral in the normalization term is the expected
number of observable exoplanets in the sample.
A further assumption that we have made in \eq{poisson-like} is that each
object is an independent sample from the rate function.
This assumption---though it is often made---cannot be correct when we consider
systems of multiple planets.
It is possible to relax this assumption (see \citealt{tremaine} for a related
technique) but this generalization is beyond the scope of the current \paper.

The main thing to note here is that $\obs{\rate}_\ratepars$ is the rate of
exoplanets that you would expect to observe taking into account the geometric
transit probability and any other detection efficiencies.
In practice, it probably makes sense to model the observable rate as
\begin{eqnarray}
\obs{\rate}_\ratepars(\entry) &=&
    \completeness(\entry)\,\rate_\ratepars(\entry)
\end{eqnarray}
where $\completeness(\entry)$ is the detection efficiency at \entry\ and
$\rate_\ratepars(\entry)$ is the object that we want to infer: the true
occurrence rate.
Up to this point, we haven't discussed any specific functional form for
$\rate_\ratepars(\entry)$ and all of this derivation is equally applicable
whether we model the occurrence rate as (for example) a broken power law or
a histogram.

@@

Now, we can substitute \eq{poisson-like} into \eq{simplemarglike} and apply
the importance sampling approximation to derive the following expression for
the marginalized likelihood
\begin{eqnarray}\eqlabel{money}
\frac{p(\{\data_k\}\,|\,\ratepars)}{p(\{\data_k\}\,|\,\interim)} &\approx&
    \exp\left(-\int \obs{\rate}_\ratepars (\entry) \dd\entry\right) \,
    \prod_{k=1}^K
    \frac{1}{N_k} \sum_{n=1}^{N_k}
    \frac{\obs{\rate}_\ratepars (\entry_k^{(n)})}
         {p(\entry_k^{(n)}\,|\,\interim)} \quad.
\end{eqnarray}
In this equation, we're making the further assumption that the catalog treated
the objects independently.
In other words, we have per-object posterior samples
\begin{eqnarray}
\entry_k^{(n)} &\sim& p(\entry_k\,|\,\data_k,\,\interim) \quad.
\end{eqnarray}
\Eq{money} is the \emph{money equation} for our method.
It lets us efficiently compute the \emph{marginalized likelihood of the entire
set of light curves for a particular occurrence rate function}.

@@

If we model the occurrence rate as a histogram with $J$ fixed bin volumes
$\binarea_j$
\begin{eqnarray}
\rate_\ratepars (\entry) &=& \left\{\begin{array}{ll}
\ratepar_1 & \entry \in \binarea_1,\\
\ratepar_2 & \entry \in \binarea_2,\\
\cdots \\
\ratepar_J & \entry \in \binarea_J,\\
0 & \mathrm{otherwise}
\end{array}\right.
\end{eqnarray}
then \eq{poisson-like} becomes
\begin{eqnarray}
\ln p(\{\entry_k\}\,|\,\ratepars) &=&
    \sum_{k=1}^K \sum_{j=1}^J \mathbf{1}[\entry_k \in
        \binarea_j]\,\ln\completeness(\entry_k)\,\ratepar_j
    -\sum_{j=1}^J\ratepar_j\,\int_{\binarea_j} \completeness(\entry)\dd\entry
\end{eqnarray}
where the indicator function $\mathbf{1}[\cdot]$ is one if $\cdot$ is true and
zero otherwise.
If we take the gradient of this function with respect to \ratepars\ and set it
equal to zero, we find the maximum likelihood result
\begin{eqnarray}
\ratepar_j &=& \frac{N_j}{\int_{\binarea_j} \completeness(\entry)\dd\entry}
\end{eqnarray}
where $N_j$ is the number of objects that fall within the bin $j$.
Since it is the maximum likelihood solution, this result is the minimum
variance estimator for the occurrence rate in the limit of negligible catalog
uncertainties.
This is \emph{almost} the same as the inverse-detection-efficiency result and
in various limiting cases, they would provide the same result.
\dfmtodo{Finish explaining why this result is better under the specified
assumptions.}

\paragraph{occurrence rate parameterization \& priors}
For the remainder of this \paper, we also model the occurrence rate as a
two-dimensional histogram with fixed logarithmic bins in period and radius.
When we include observational uncertainties---using \eq{money}---the maximum
likelihood result is no longer analytic.
Therefore, if we want to compute the ``best-fit'' occurrence rate, we can use a
standard non-linear optimization algorithm.
In the regions of parameter space that we tend to care about, the completeness
is low and there are only a few observations with large uncertainties.
In this case, we're especially interested in probabilistic constraints on the
occurrence rate; not just the best-fit model.
To do this, we apply a prior\footnote{This is a slight misnomer because
we also we also marginalize over the parameters of this model.} $p(\ratepars)$
on the occurrence rate parameters and generate samples from the posterior
probability
\begin{eqnarray}\eqlabel{posterior}
p(\ratepars\,|\,\{\data_k\}) &\propto&
    p(\ratepars)\,p(\{\data_k\}\,|\,\ratepars)
\end{eqnarray}
using Markov chain Monte Carlo (MCMC).

Given our model of the rate function as a collection of bins, we choose to
model the prior as a Gaussian process (\citealt{gp,gibson-gp}) on the
logarithmic bin heights.
This model encodes our prior belief that, on the grid scale that we consider,
the occurrence rate should be smooth but it is otherwise very flexible about
the form of the function.
Mathematically, the prior is written as
\begin{eqnarray}
p(\ratepars) &=& p(\ratepars\,|\,\mean,\,\smooth) \\
&=& \mathcal{N} \left[\ln\ratepars;\,\mean\,\bvec{I},\,
\cov(\{\binarea_j\},\,\smoothpars)\right]
\end{eqnarray}
where $\mathcal{N}(\cdot;\,\mean\,\bvec{I},\,\cov)$ is a $J$-dimensional
Gaussian with a constant mean \mean\ and covariance matrix \cov\ that depends
on the bin locations $\{\binarea_j\}$ and a set of hyperparameters
$\smoothpars = (\smooth_0,\,\smooth_\period,\,\smooth_\radius)$.
The covariance function that we use is an anisotropic, axis-aligned
exponential-squared kernel so elements of the matrix are
\begin{eqnarray}
\cov_{ij} &=& \smooth_0\,\exp\left(-\frac{1}{2}\,
    [\binarea_i-\binarea_j]^\mathrm{T}\,\Sigma^{-1}\,[\binarea_i-\binarea_j]
\right)
\end{eqnarray}
where $\Sigma^{-1}$ is the diagonal matrix
\begin{eqnarray}
\Sigma^{-1} &=& \left(\begin{array}{cc}
1/\smooth_\period & 0 \\
0 & 1/\smooth_\radius
\end{array}\right) \quad.
\end{eqnarray}

@@

In order to get a feeling for the constraints provided by our method and to
explore any biases introduced by ignoring the observational uncertainties, we
start by ``observing'' a few synthetic catalogs from qualitatively different
known occurrence rate functions.
For each of these simulations, we'll take the completeness function computed
by \citet{petigura} as given.
In general, \eq{poisson-like} can be sampled using a procedure called thinning
(\citealt{poisson}) but for our purposes, we'll simply consider a piecewise
constant rate function evaluated on a fine grid in log-period and log-radius.
For this discrete function, the generative procedure is simple;
\begin{enumerate}
{\item loop over each grid cell $i$,}
{\item draw Poisson random integer $K_i\sim\mathrm{Poissson}(\obs{\rate}_i)$
with the observable rate evaluated in the cell, and}
{\item distribute $K_i$ catalog entries in the cell randomly.}
\end{enumerate}
We then choose fractional observational uncertainties on the radii from the
\citet{petigura} catalog and apply them to the true catalog as Gaussian noise.

We generate synthetic catalogs from two qualitatively different rate
functions.
Both distributions are generated by a separable model
\begin{eqnarray}
\rate_\ratepars (\ln\period,\,\ln\radius) &=&
    \rate_\ratepars^{(\period)}(\ln\period)\,
    \rate_\ratepars^{(\radius)}(\ln\radius)
\end{eqnarray}
but fit using the full general model.
The first catalog---\modela--- is generated assuming a smooth occurrence
surface where both distributions are broken power laws.
The second---\modelb---is designed to be exactly the distribution inferred by
\citet{petigura} in the range that they considered and then smoothly
extrapolated outside that range.
The catalogs generated from these two models are shown in \fig{smooth-results}
and \fig{simulation-results}, respectively and the data are available online
at \dfmtodo{ADD SOME URL}.

For each catalog, we directly apply the inverse-detection-efficiency procedure
as implemented by \citealt{petigura}\footnote{Our implementation reproduces
their results when applied to the published catalog.} and our probabilistic
method, marginalizing over the hyperparameters of the Gaussian process
regularization.
\Fig{smooth-results} and \fig{simulation-results} show the results of this
analysis in both cases.
In particular, the side panels compare the marginalized occurrence rate
functions in period and radius to the true functions that were used to
generate the catalogs.
\Fig{smooth-results} shows that even if the true occurrence rate is a smooth
function, the rate inferred by the inverse-detection-efficiency method can
appear to have sharp features.
In this first example---where the true distribution is well described by our
Gaussian process model---the probabilistic inference of the occurrence rate is
both more precise and accurate.

In the second example, the true rate function includes a sharp feature chosen
to reproduce the result published by \citet{petigura}.
In this case, \fig{simulation-results} shows that the probabilistic
constraints on the occurrence rate are less precise \emph{but more accurate}
than results using the inverse-detection-efficiency method.
At this point, the distributions inferred by these two different techniques
don't seem drastically inconsistent but the difference are most obvious as the
survey completeness goes to zero: exactly the region of interest.

\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{figures/smooth/results.pdf}
\end{center}
\caption{%
{\bf Simulated data}.
Inferences about the occurrence rate function based on the simulated catalog
\modela.
\emph{Center:} the points with error bars show the exoplanet candidates in the
simulated incomplete catalog, the contours show the survey completeness
function (\citealt{petigura}), and the grayscale shows the median posterior
occurrence surface.
\emph{Top and left:} the red dashed line shows the \emph{true} distribution
that was used to generate the catalog, the points with error bars show the
results of the inverse-detection-efficiency procedure, and the histograms are
posterior samples from the marginalized occurrence rate as inferred by our
method.
\figlabel{smooth-results}}
\end{figure}

@@

Having developed this probabilistic framework for occurrence rate inferences
and demonstrating that it produces reasonable results when applied to
simulated datasets, we now turn to real data.
As described in \sect{data}, we will use the catalog of small exoplanet
candidates orbiting Sun-like stars published by \citet{petigura}.
This is a great test case because those authors empirically measured the
detection efficiency of their pipeline as a function of the parameters of
interest.
